A Case Study on Abstractions to Converge High-Throughput and High-Performance Computing

### 2017-03-21 Experiments.tex
New iteration experiments

In this section we present experiments that aim to determine the performances of the NGE and to show how it provides a negligible overhead while introducing new functionalities.

Experiments consist in running AthenaMP instances by using NGE pilots where each AthenaMP simulates a pre-determined number of events taken from ATLAS workload.	
We present four groups of experiments in which we test the NGE for weak scalability, weak scalability with resubmission, strong scalability and heterogeous workloads.
For each group we vary the size of the pilots and the number of AthenaMP that are sent for execution. 
We collect data about the execution time of the pilots and the AthenaMP that are executed within them by considering all stages that these entities pass through during the execution.
Experiments have been performed out of production by using NGE's pilots on TITAN's batch queue. 
Because of TITAN's batch queue, the turnaround time of each run of our experiments is dominated by the time spent on TITAN's queue. Since we are interested only in the performances of the NGE, we do not provide data about queue time.

For the experiment, AthenaMP has been set to use all the 16 cores present on TITAN's nodes.

@@

Tasks do not experience queue within the pilot since there is one node for each ATLAS simulation. Therefore, delays are consequence of only three factors: i) the bootstrapping of the pilot on the nodes; ii) the manager that has to dispatch tasks to the pilot; iii) the time that the agent requires to spawn all of tasks on the nodes.

@@

Figure \ref{fig:weakScal1} depicts the average pilot duration, the average execution time of each tasks and the average time required to send a tasks for execution.

\begin{figure}[!htb]
        \includegraphics[width=0.5\textwidth]{./figures/NGE/Weak.pdf}
    \caption{Average pilot execution time, average tasks execution time and average launching time for pilot sizes: 250, 500, 1000 and 2000.}
\label{fig:weakScal1}
\end{figure}
\aanote{TO be commented as soon as the figure becomes definitive. Note that It would be better if we show the ratio between pilot duration and execution time}
\subsubsection{Weak scalability with resubmission }
This experiment is similar to the one presented above but in this case we want to test also the impact of submitting new simulations in place of those that end. Thus, we send to the pilot a number of tasks that is five times larger than the number of nodes. The reason behind this experiment is to stress the manager and the agent by pushing new tasks while others are ending their execution in such a way to increase the concurrency of these two components.

In order to perform the experiments in a reasonable amount of time, we reduced  the number of events that are simulated by Athena-MP to sixteen. This allows us to complete a task in ~25 minutes only. Note that the number of events has been chosen in such a way that all the cores of a node run one event.

We tested different pilot sizes, i.e. : 250, 500, 1000 and 2000. For all of them the walltime was 3 hours.

Figure \ref{fig:weakScal2} \aanote{(DRAFT)}  depicts the average pilot duration, the average execution time of each tasks and the average time required to send a tasks for execution.

\begin{figure}[!htb]
        \includegraphics[width=0.48\textwidth]{./figures/draft/AvgT.pdf}
    \caption{Average pilot execution time, average tasks execution time and average launching time for pilot sizes: 250, 500, 1000 and 2000.}
\label{fig:weakScal2}
\end{figure}

\subsubsection{Strong scalability}

@@

>>>>

In this section we present experiments that aim to determine the performances of the NGE and to show how it provides a minimal overhead while introducing new functionalities.

Experiments consist in running AthenaMP instances by using NGE pilots where each AthenaMP simulates a pre-determined number of events taken from ATLAS workload.	
We present four groups of experiments in which we test the NGE for weak scalability, weak scalability with resubmission, strong scalability and heterogeous workloads.

We collect data about the execution time of the pilots and the AthenaMP that are executed within them by considering all stages that these entities pass through during the execution.
Experiments have been performed out of production by using NGE's pilots on TITAN's batch queue. 
Because of TITAN's batch queue, the turnaround time of each run of our experiments is dominated by the time spent on TITAN's queue. Since we are interested only in the performances of the NGE, we removed queue time from our statistics.

All the experiments have been performed by setting AthenaMP in such a way that it uses all the 16 cores present on TITAN's nodes.

@@

Tasks do not experience queue within the pilot since there is one node for each AthenaMP instance. Therefore, delays are consequence of only three factors: i) the bootstrapping of the pilot on the nodes; ii) the manager, as defined in Section \ref{}, that has to dispatch tasks to the pilot; iii) the time that the agent requires to spawn all the tasks on the nodes.

@@

Figure \ref{fig:weakScal1a} depicts the average pilot duration and the average execution time of AthenaMP as function of the pilot size.

\begin{figure}[!htb]
        \includegraphics[width=0.5\textwidth]{./figures/NGE/weakET1.pdf}
    \caption{Average pilot execution time against average AthenaMP execution times for pilot sizes: 250, 500, 1000 and 2000.}
\label{fig:weakScal1a}
\end{figure}

On the one side, we can observe that the average pilot duration is insensitive to the pilot size. This indicates that, on average, the overhead introduced by the pilot does not grow significatively with the size of the pilot.
On the other side, we can notice that the average execution time of AthenaMP is smaller than the average pilot duration. This indicates that the overhead of the pilot is not null and can be quantified in ~5 minutes.

\aanote{I think we should say something about the execution time of AthenaMP that is smaller than the others.}

In Figure \ref{fig:weakScal1b}, we provide the time required by the pilot to start running AthenaMP instances.  
We can notice that the larger is the number of units and the larger is the time required by the pilot to start running them. However, we can observe that the time does not grow linearly with the pilot size.

\begin{figure}[!htb]
        \includegraphics[width=0.5\textwidth]{./figures/NGE/weakStart1.pdf}
    \caption{Average time to start running AthenaMP for pilot sizes: 250, 500, 1000 and 2000.}
\label{fig:weakScal1b}
\end{figure}

\subsubsection{Weak scalability with resubmission }
This experiment is similar to the one presented above but in this case we want to test also the impact of submitting new AthenaMP instances in place of those that end. The reason behind this experiment is to stress pilot's components by pushing new tasks while others are ending their execution.

In order to perform the experiments in less than two hours per run, we decided to send for execution a number of AthenaMP instances equal to five times the number of nodes and we reduced  the number of events that are simulated by each Athena-MP to sixteen\footnote{Note that the number of events has been chosen in such a way that all the cores of a node run one event.}. This allows us to complete a single AthenaMP in ~20 minutes only.

We tested different pilot sizes, i.e. : 256, 512, 1024 and 2048. For all of them the walltime was 3 hours.

Figure \ref{fig:weakScal2a} depicts the average pilot duration and  the average execution time of AthenaMP.

\begin{figure}[!htb]
        \includegraphics[width=0.5\textwidth]{./figures/NGE/weakET2.pdf}
    \caption{Average pilot execution time against average AthenaMP execution times  for pilot sizes: 256, 512, 1024 and 2048.}
\label{fig:weakScal2a}
\end{figure}
\begin{figure}[!htb]
        \includegraphics[width=0.5\textwidth]{./figures/NGE/weakStart2.pdf}
    \caption{Average time to start running AthenaMP for pilot sizes: 256, 512, 1024 and 2048.}
\label{fig:weakScal2b}
\end{figure}
\subsubsection{Strong scalability}
This experiments aim to test strong scalability. Therefore, we run the same amount of tasks for different pilot sizes. 
We used a number of AthenaMP instances equal to 2048 and tested pilots with 256, 512, 1024 and 2048 nodes. 
Each AthenaMP instance simulates sixteen events as the previous experiment.

Figure \ref{fig:strongScala} depicts the average pilot duration and  the average execution time of AthenaMP.

%\begin{figure}[!htb]
%        \includegraphics[width=0.5\textwidth]{./figures/NGE/strongET.pdf}
%    \caption{Average pilot execution time against average AthenaMP execution times  for pilot sizes: 256, 512, 1024 and 2048.}
%\label{fig:strongScala}
%\end{figure}
%\begin{figure}[!htb]
%        \includegraphics[width=0.5\textwidth]{./figures/NGE/strongStart.pdf}
%    \caption{Average time to start running AthenaMP for pilot sizes: 256, 512, 1024 and 2048.}
%\label{fig:strongScalb}
%\end{figure}

@@

The last experiment provides a proof of concept about the ability of the NGE to execute heterogeneous workload.
In particular, we coupled the execution of AthenaMP with the execution of Gromacs to simulate molecular dynamics. 
We performed the experiment by executing at first one AthenaMP per node and then, submitting a Gromacs simulation per core. Each Gromacs simulation requires ~20 minutes.
We tested the following pilot sizes : 8, 32, 64.
